Maximum Likelihood

### 2017-03-22 sources/Math/MaximumLikelihood.rmd
maximum likelihood pdf

>>>>

For the likelihood of two points we are going to use the same method to find the best Gaussian distribution to explain both observations together.

This means:

L(theta|x) = L(theta|x1,x2) = p(x1,x2|theta)

To simplify let start optimizing this function for u first. So we have:

p(x1,x2|u) = p(x1|x2)p(x2) = p(x2|x1)p(x1)

We generraly assume that both observations are independent because in this way we have that

p(x2|x1) = p(x2)  
p(x1|x2) = p(x1)

so

p(x1,x2|u) = p(x1)p(x2)

This means that the we must optimize the product of the probability density function.

First let plot this product value for various u values. With this plot we can clearly see that the function is optimized when the product is maximized.

```{R}
A <- -0.25  
B <- -1
obs <- c(A,B)
r <- seq(-5,5, by = 0.1)
plot(r,sapply(r, function(x) {dnorm(A, mean = x, sd = 1)}), col = "red")
points(r,sapply(r, function(x) {dnorm(B, mean = x, sd = 1)}), col = "blue")
points(r,sapply(r, function(x) {dnorm(A, mean = x, sd = 1)*dnorm(B, mean = x, sd = 1)}))
arrows(A,0,A,dnorm(A,A,1), col = "red")
arrows(B,0,B,dnorm(B,B,1), col = "blue")
```

So, first lets try to optimize the function numeracally:

```{R}
A <- -0.25  
B <- -1
f <- function (x) {dnorm(A, mean = x, sd = 1)*dnorm(B, mean = x, sd = 1)}
xmax <- optimize(f, c(-5, 5), tol = 0.0001, maximum = TRUE)

A <- -0.25  
B <- -1
obs <- c(A,B)
r <- seq(-5,5, by = 0.1)
plot(r,sapply(r, function(x) {dnorm(A, mean = x, sd = 1)}), col = "red")
points(r,sapply(r, function(x) {dnorm(B, mean = x, sd = 1)}), col = "blue")
points(r,sapply(r, function(x) {dnorm(A, mean = x, sd = 1)*dnorm(B, mean = x, sd = 1)}))
arrows(A,0,A,dnorm(A,A,1), col = "red")
arrows(B,0,B,dnorm(B,B,1), col = "blue")
arrows(xmax$maximum,0,xmax$maximum,xmax$objective, col = "green")
```

So now we now that:
L(theta|x)    
  = L(theta|x1,x2)  
  = p(x1,x2|theta) (for simplicity theta={u})   
  = p(x1|x2)p(x2) for Gaussian(u,sd = 1)  
  = p(x1)p(x2) for Gaussian(u,sd = 1)

argmax(u) in p(x1)p(x2) for Gaussian(u,sd = 1)  
  maximum at `r xmax$maximum`

so the best distribution is: Gaussian(u = `r xmax$maximum`, sd = 1)
